\documentclass[GTS.tex]{subfiles}

\addcontentsline{toc}{chapter}{Bibliografía}

\begin{document}

\renewcommand\chaptername{\Huge Tema}

\titleformat{\chapter}[display]
    {\normalfont\huge\bfseries}{\chaptertitlename\ \thechapter}{10pt}{\Huge}
\titlespacing*{\chapter}{0pt}{-1cm}{10pt}

\begin{thebibliography}{}
\bibitem{Zeeman} E. C. Zeeman, \textit{An Introduction to Topology. The Clasification theorem for Surfaces.}\\ \url{http://www.maths.ed.ac.uk/~aar/surgery/zeeman.pdf}
\bibitem{Hatcher} A. Hatcher, \textit{Algebraic Topology}. Cambridge University Press, 2002. \\ \url{https://www.math.cornell.edu/~hatcher/AT/AT.pdf}
\bibitem{Moise} E. E. Moise, \textit{Geometric Topology in Dimensions 2 and 3. } Springer-Verlag, 1977.
\bibitem{Doyle} P. H. Doyle \& D. A. Moran, \textit{A Short Proof that Compact 2-Manifolds Can Be Triangulated.} Inventiones math. 5 (1968), 160-162. \\	    	\url{http://www.maths.ed.ac.uk/~aar/papers/doylemoran.pdf}
\bibitem{Bloch} E. D. Bloch, \textit{A First Course in Geometric Topology and Differential Geometry.} Birkhäuser, 1977.
\bibitem{Mohar} B. Mohar \& C. Thomassen, \textit{Graphs on Surfaces}. Johns Hopkins University Press, 2001.
\bibitem{Xu} J. Gallier \& D. Xu, \textit{A Guide to the Classification Theorem for Compact Surfaces}. Springer, 2013.
\bibitem{} A. Cañas, \textit{El grupo fundamental como clasificador de superficies compactas.} Trabajo de Fin de Grado, Universidad de La Rioja, 2013. \url{http://biblioteca.unirioja.es/tfe_e/TFE000323.pdf}
\bibitem{} M. A. Armstrong, \textit{Topología Básica}. Reverte, 1987.
\bibitem{} J. R. Munkres, \textit{Topología.} Pearson Educación, 2002.\\ \url{https://psm73.files.wordpress.com/2009/11/topologia-munkres.pdf}
\end{thebibliography}
\end{document}